Air Quality Data - PM2.5 Components

### 2021-01-03 code/data-processors/air-data-processor.Rmd
create 6, 12, 24 mile buffers around sites

aqs.sites <- read.csv(unz(temp,"aqs_sites.csv")) %>%
  filter(
    #remove monitors closed before 2000
    year(as.Date(Site.Closed.Date)) >= 2000 | Site.Closed.Date == "",
    !State.Name %in% c("Alaska","Hawaii","Puerto Rico","Virgin Islands",
                       "Guam", "Country Of Mexico","Canada") #remove non-CONUS states
    ) %>%  
  transmute(
    Site.ID = as.factor(sprintf("%02d-%03d-%04d", #unique ID for each Site: ss-ccc-nnnn
                                as.numeric(as.character(State.Code)),
                                County.Code, Site.Number)), 
    State.Code = as.factor(sprintf("%02d", as.numeric(as.character(State.Code)))),
    County.Code = as.factor(sprintf("%03d", County.Code)),
    Latitude,
    Longitude, 
    State = state.abb[match(State.Name,state.name)],
    State = as.factor(ifelse(State.Name == "District Of Columbia","DC", State)),
    Land.Use,
    Location.Setting = as.factor(ifelse(Location.Setting == "URBAN AND CENTER CITY",
                                        "URBAN",
                                        as.character(Location.Setting)))
    ) %>% 
  merge(state.region[, c("State", "Region.IV")])

save(aqs.sites, file = here('data', 'processed', 'aqs-sites.RDa'))

aqs.sites %>% 
  group_by(Location.Setting) %>% 
  summarise(n = n())
```

@@

```{r check carbon}
### EC
carbon.ec <- dt.spec.raw %>% 
  select(Site.ID, POC, Monitor.ID, Parameter.Symbol, Date, Arithmetic.Mean) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(Arithmetic.Mean)) %>% 
  arrange(Site.ID, POC, Date) %>%
  select(Site.ID, POC, Monitor.ID, EC.TOR, EC.TOT)

ec.c<- carbon.ec[complete.cases(carbon.ec),] #405 simultaneous TOR & TOT
ec.c %>% group_by(Site.ID) %>% 
  summarise(n = n()) # 11 Sites ~40 each

@@

### OC
carbon.ec <- dt.spec.raw %>% 
  select(Site.ID, POC, Monitor.ID, Parameter.Symbol, Date, Arithmetic.Mean) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(Arithmetic.Mean)) %>% 
  arrange(Site.ID, POC, Date.Local) %>%
  select(Site.ID, POC, Monitor.ID, OC.TOR, OC.TOT)

oc.c<- carbon.oc[complete.cases(carbon.oc),] #405 simultaneous TOR & TOT
oc.c %>% group_by(Site.ID) %>% 
  summarise(n = n())

@@

site.comp <- dt.comp.1yr %>% distinct(Site.ID) %>%  #574 -> 486 -> 346 -> 346
  merge(aqs.sites)
save(site.comp, file = here('data', 'processed', 'site-comp.RDa'))

site.comp %>% 
  group_by(Location.Setting) %>% 
  summarise(n = n())
#Location.Setting  n
#                  2			
#RURAL	           202			
#SUBURBAN	         69			
#URBAN	           73	
```

>>>>

dt.aqs.sites <- read.csv(unz(temp,"aqs_sites.csv")) %>%
  filter(
    #remove monitors closed before 2000
    year(as.Date(Site.Closed.Date)) >= 2000 | Site.Closed.Date == "",
    !State.Name %in% c("Alaska","Hawaii","Puerto Rico","Virgin Islands",
                       "Guam", "Country Of Mexico","Canada") #remove non-CONUS states
    ) %>%  
  transmute(
    Site.ID = as.factor(sprintf("%02d-%03d-%04d", #unique ID for each Site: ss-ccc-nnnn
                                as.numeric(as.character(State.Code)),
                                County.Code, Site.Number)), 
    State.Code = as.factor(sprintf("%02d", as.numeric(as.character(State.Code)))),
    County.Code = as.factor(sprintf("%03d", County.Code)),
    Latitude,
    Longitude, 
    State = state.abb[match(State.Name,state.name)],
    State = as.factor(ifelse(State.Name == "District Of Columbia","DC", State)),
    Land.Use,
    Location.Setting = as.factor(ifelse(Location.Setting == "URBAN AND CENTER CITY",
                                        "URBAN",
                                        as.character(Location.Setting)))
    ) %>% 
  merge(dt.state.region[, c("State", "Region.IV")])

save(dt.aqs.sites, file = here('data', 'processed', 'aqs-sites.RDa'))

dt.aqs.sites %>% 
  group_by(Location.Setting) %>% 
  summarise(n = n())
```

@@

```{r check carbon}
### EC
dt.carbon.ec <- dt.spec.raw %>% 
  select(Site.ID, POC, Monitor.ID, Parameter.Symbol, Date, Arithmetic.Mean) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(Arithmetic.Mean)) %>% 
  arrange(Site.ID, POC, Date) %>%
  select(Site.ID, POC, Monitor.ID, EC.TOR, EC.TOT)

ec.c<- dt.carbon.ec[complete.cases(dt.carbon.ec),] #405 simultaneous TOR & TOT
ec.c %>% group_by(Site.ID) %>% 
  summarise(n = n()) # 11 Sites ~40 each

@@

### OC
dt.carbon.oc <- dt.spec.raw %>% 
  select(Site.ID, POC, Monitor.ID, Parameter.Symbol, Date, Arithmetic.Mean) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(Arithmetic.Mean)) %>% 
  arrange(Site.ID, POC, Date.Local) %>%
  select(Site.ID, POC, Monitor.ID, OC.TOR, OC.TOT)

oc.c<- dt.carbon.oc[complete.cases(dt.carbon.oc),] #405 simultaneous TOR & TOT
oc.c %>% group_by(Site.ID) %>% 
  summarise(n = n())

@@

dt.site.comp <- dt.comp.1yr %>% distinct(Site.ID) %>%  #574 -> 486 -> 346 -> 346
  merge(dt.aqs.sites)
save(dt.site.comp, file = here('data', 'processed', 'site-comp.RDa'))

dt.site.comp %>% 
  group_by(Location.Setting) %>% 
  summarise(n = n())
#Location.Setting  n
#                  2			
#RURAL	           202			
#SUBURBAN	         69			
#URBAN	           73	
```

```{r Zipcodes in x-mile bufferzone of Monitoring Sites}
load(here('data', 'processed', 'comp-1yr.RDa'))

dt.comp <- dt.comp.1yr %>% 
  select(Site.ID, Year, Month) %>% 
  mutate(psuedo = 100)

# joining zipcodes to monitoring sites
dt.comp <- dt.comp %>% 
  inner_join(dt.site.zip, by = c("Site.ID", "Year")) %>% 
  arrange(Zip.Code, Year, Dist)

# pick the site with shorter distance to zipcode centroeid
dt.site.dist.shorter <- dt.comp %>% 
  group_by(Zip.Code, Year) %>% 
  filter(Dist == min (Dist)) %>% 
  arrange(Zip.Code, Year) %>% 
  as.data.frame()

# 6-mile bufferzone around monitoring site
dt.comp.site.zip.bz6 <- dt.site.dist.shorter %>% 
  filter(Dist <= 6) %>% 
  select(Zip.Code, Year, Month, Site.ID, psuedo)

save(dt.comp.site.zip.bz6, file = here('data', 'processed', 'comp-site-zip-bz6.RDa'))
write.csv(dt.comp.site.zip.bz6, file = here('data', 'processed', 'comp-site-zip-bz6.csv'))

# Count number of zipcodes and sites
dt.comp.site.zip.bz6 %>% distinct(Zip.Code) %>% nrow() #4351
dt.comp.site.zip.bz6 %>% distinct(Site.ID) %>% nrow() #277

# 12-mile bufferzone around monitoring site
dt.comp.site.zip.bz12  <- dt.site.dist.shorter %>% 
  filter(Dist <= 12) %>% 
  select(Zip.Code, Year, Month, Site.ID, psuedo)

save(dt.comp.site.zip.bz12, file = here('data', 'processed', 'comp-site-zip-bz12.RDa'))
write.csv(dt.comp.site.zip.bz12, file = here('data', 'processed', 'comp-site-zip-bz12.csv'))

# Count number of zipcodes and sites
dt.comp.site.zip.bz12 %>% distinct(Zip.Code) %>% nrow() #8454
dt.comp.site.zip.bz12 %>% distinct(Site.ID) %>% nrow() #328

# 24-mile bufferzone around monitoring site
dt.comp.site.zip.bz24  <- dt.site.dist.shorter %>% 
  filter(Dist <= 24) %>% 
  select(Zip.Code, Year, Month, Site.ID, psuedo)

save(dt.comp.site.zip.bz24, file = here('data', 'processed', 'comp-site-zip-bz24.RDa'))
write.csv(dt.comp.site.zip.bz24, file = here('data', 'processed', 'comp-site-zip-bz24.csv'))

# Count number of zipcodes and sites
dt.comp.site.zip.bz24 %>% distinct(Zip.Code) %>% nrow() #15856
dt.comp.site.zip.bz24 %>% distinct(Site.ID) %>% nrow() #341
```
